Errata

### 2025-04-06 tex/Manuscript-5.tex
Update Manuscript-5.tex: Refactor table structure for clarity and improve formatting
Add auxiliary files for Methodology and Results_Discussion chapters
- Created Methodology.aux with references, labels, and citations for chapter 2.
- Created Results_Discussion.aux with references, labels, and citations for chapter 3.
- Updated table and figure references for improved organization and clarity.

    \vspace{0.5cm}

    Kristina Witzgall\textsuperscript{1}, Franziska Steiner\textsuperscript{1}, Benjamin D. Hesse\textsuperscript{2,3}, Nicolás Riveras-Muñoz\textsuperscript{4}, Victoria Rodríguez\textsuperscript{5}, Pedro Teixeira\textsuperscript{1}, Li, M.\textsuperscript{6}, Rómulo Oses\textsuperscript{7}, Oscar Seguel\textsuperscript{8}, Steffen Seitz\textsuperscript{4}, Dirk Wagner\textsuperscript{5,6}, Thomas Scholten\textsuperscript{4}, Franz Buegger\textsuperscript{9} Gerrit Angst\textsuperscript{7,11,12}, and Carsten W. Mueller\textsuperscript{1,13,14}

    \vspace{0.2cm}
  \end{center}

@@

A critical aspect of the plant-associated contribution to these soil processes occurs at the root-soil interface, i.e., within the rhizosphere. As plants grow, their roots physically alter soil minerals, and root-associated microorganisms release weathering agents (e.g., organic acids; \citep{Uroz2009}) that mobilize essential soil nutrients \citep{Gregory2022}. 
In addition, the release of rhizodeposits at the root-soil interface represents a major pathway for the build-up of rather persistent mineral-associated OM (MAOM) in soil \citep{Sokol2019, Villarino2021}. Moreover, the increased abundance of labile C in the rhizosphere stimulates the growth and activity of microorganisms. This in turn drives the formation of aggregates via microbially derived binding agents (i.e., extracellular polymeric substances; EPS) or via the enmeshment facilitated by fungal hyphae \citep{Chenu2011, Costa2018}. In addition, root exudates (e.g., root mucilage or polysaccharide exudates) can either adhere directly to mineral surfaces or function as glue-like binding agents, binding soil particles together, forming MAOM \citep{Roetzer2023} and driving soil aggregation \citep{Baumert2018}. As such, roots substantially shape the chemical, physical, and biological nature of the soil at the root-soil interface, and the rhizosphere is widely recognized as a dynamic system in both space and time \citep{Jones2009, Hinsinger2009, Kuzyakov2019}. However, relatively little is known about the temporal changes in processes occurring in the rhizosphere, as well as potential long-term effects of the living root system that may persist beyond the lifetime of the roots \citep{Oliver2021}. These effects, so-called 'root legacy effects', might have implications for soil physical, chemical, and biological parameters that can remain in the soil after plant death and thus shape the subsequent detritusphere \citep{Wurst2015}.

@@

Soil was sampled from five subplots on south-facing top slopes within the Santa Gracia Natural Reserve, located in the Coquimbo Region in Chile (\ang{29;44;26.3}S, \ang{71;09;01.2}W) with mean annual precipitation of \SI{89}{\milli\metre\,\year^{-1}} and mean annual air temperature of \SI{13.7}{\degreeCelsius} \citep{MinisterioObrasPublicas2017}. All five subplots were classified as Cambisol according to the WRB \citep{IUSSWorkingGroupWRB2015}. Vegetation was dominated by shrubs (\SIrange{30}{40}{\percent} coverage; including \textit{Proustia cuneifolia}, \textit{Balbisia peduncularis}, \textit{Cordia decandra} and \textit{Baccharis paniculatum}), classified as 'Interior Mediterranean desert scrub' \citep{Luebert2006}, and heavily disturbed by grazing animals \citep{Oeser2018}. The sampling site was selected as a suitable representation of an initial dryland soil with low vegetation cover, low organic C content, shallow and weakly structured A horizon \citep{Bernhard2018}. Topsoil and subsoil material was collected and sieved (\SI{< 2}{\milli\metre}) from the profile walls of A and B horizons (\SIrange{0}{3}{\centi\metre} and \SIrange{5}{30}{\centi\metre}, respectively), representing soils at different stages of development. This allowed us to directly compare topsoil and subsoil, but also to use the subsoil as a model system for initial soil formation via plant-soil interactions in a saprolitic material.

@@

The topsoil and subsoil material from the five subplots was homogenized and cleared of plant residues and seeds to prevent unwanted germination during incubation. The two soil materials (topsoil and subsoil) were weighed separately into pots (\SI{350}{\centi\metre\cubed}) aiming to resemble bulk density according to field conditions (\SIrange{1.6}{1.7}{\gram\per\centi\metre\cubed}; \citep{Bernhard2018}). Pre-germinated seeds (\textit{Helenium aromaticum} (Hook.) L. H. Bailey) were sown on half of the pots, while the other half was left free of plants and hereafter referred to as controls. After the initial incubation period of 70 days under greenhouse conditions ('rhizosphere phase'), half of the samples were harvested, while the remaining half were placed in complete darkness and incubated at room temperature for 100 days ('detritusphere phase'). The samples were left completely undisturbed during decomposition, meaning that above-ground biomass was left to decompose on the soil surface. During both incubation phases, the samples received \SI{6}{\milli\liter} of rainwater per day, without correcting for potential differences in soil moisture, e.g. due to differences in plant water use between treatments.

@@

We tested all parameters for homoscedasticity using Levene's test and the residuals of each model were tested for Normality using QQ-Plots and the Shapiro-Wilk test. In cases where the assumption of normality or homoscedasticity was not met, the data was appropriately transformed. To evaluate differences between the data presented in \ref{tab:M5-T1}, we conducted a one-way analysis of variance (ANOVA) with Tukey's honestly significant differences (HSD) as a post-hoc test. The remaining data was analyzed using mixed-effect models (package: \texttt{nlme}) using soil material (topsoil vs subsoil) and phase (rhizosphere phase vs detritusphere phase) as fixed factors and the ID as a random effect. All statistical testing was conducted in R (version 4.2.1; \citep{RCoreTeam2008}) using RStudio (Version 2022.07.2, R Core Team, Vienna, Austria; \citep{RStudioTeam2015}). Further, data shown in text and tables are given as the mean values and standard deviations.

\begin{sidewaystable}[htbp]

@@

  \begin{tabular}{llcccccccc}

@@

  \multirow{3}{*}{Category} & \multirow{3}{*}{Property} & \multicolumn{4}{c}{\textbf{Topsoil}} & \multicolumn{4}{c}{\textbf{Subsoil}} \\
  \cmidrule(lr){3-6} \cmidrule(lr){7-10}
  & & \multicolumn{2}{c}{Rhizosphere} & \multicolumn{2}{c}{Detritusphere} & \multicolumn{2}{c}{Rhizosphere} & \multicolumn{2}{c}{Detritusphere} \\
  \cmidrule(lr){3-4} \cmidrule(lr){5-6} \cmidrule(lr){7-8} \cmidrule(lr){9-10}
  & & Above & Below & Above & Below & Above & Below & Above & Below \\

@@

  \multirow{8}{*}{Basic properties}
  & C (mg g$^{-1}$) & 4.06 $\pm$ 0.20 & 3.70 $\pm$ 0.08 & 3.49 $\pm$ 0.47 & 3.97 $\pm$ 0.74 & 3.16 $\pm$ 0.50 & 3.70 $\pm$ 0.09 & 3.23 $\pm$ n.a. & 3.36 $\pm$ 1.43 \\
  & N (mg g$^{-1}$) & 0.17 $\pm$ 0.05** & 0.11 $\pm$ 0.02** & 0.25 $\pm$ 0.06** & 0.11 $\pm$ 0.02** & 0.09 $\pm$ 0.01 & 0.07 $\pm$ 0.00 & 0.11 $\pm$ n.a. & 0.10 $\pm$ 0.04 \\
  & C:N & 25.61 $\pm$ 7.77** & 34.43 $\pm$ 3.43** & 14.67 $\pm$ 3.99*** & 37.26 $\pm$ 8.89 & 33.58 $\pm$ 2.90 & 53.54 $\pm$ 5.27*** & 28.70 $\pm$ n.a. & 32.37 $\pm$ -- \\
  & Biomass (g) & 0.47 $\pm$ 0.17*** & 0.44 $\pm$ 0.13*** & 0.11 $\pm$ 0.07*** & 0.03 $\pm$ 0.02*** & 0.25 $\pm$ 0.09*** & 0.29 $\pm$ 0.07*** & 0.04 $\pm$ n.a. & 0.02 $\pm$ 0.00*** \\
  & Total biomass (g) & 0.90 $\pm$ 0.22*** & 0.13 $\pm$ 0.09*** & 0.54 $\pm$ 0.11*** & 0.05 $\pm$ 0.07*** & \multicolumn{4}{c}{--} \\
  & Root:shoot & 1.07 $\pm$ 0.48** & 0.32 $\pm$ 0.20** & 1.27 $\pm$ 0.37 & 0.24 $\pm$ n.a. & \multicolumn{4}{c}{--} \\
  & C (mg) & 1.91 $\pm$ 0.78*** & 1.62 $\pm$ 0.51*** & 0.38 $\pm$ 0.29*** & 0.12 $\pm$ 0.09*** & 0.75 $\pm$ 0.21 & 1.07 $\pm$ 0.12*** & 0.36 $\pm$ n.a. & 0.06 $\pm$ 0.04*** \\
  & N (mg) & 0.07 $\pm$ 0.02*** & 0.05 $\pm$ 0.01*** & 0.03 $\pm$ 0.02*** & 0.00 $\pm$ 0.01*** & 0.02 $\pm$ 0.01 & 0.02 $\pm$ 0.00 & 0.01 $\pm$ n.a. & 0.00 $\pm$ 0.00 \\
  & $\delta^{15}$N ($\permil$ air N$_2$) & 4.96 $\pm$ 0.78 & 4.14 $\pm$ 0.59$^{\bullet}$ & 4.29 $\pm$ 0.71 & 3.22 $\pm$ 0.81$^{\bullet}$ & 4.66 $\pm$ 0.61 & 4.41 $\pm$ 0.45*** & 3.94 $\pm$ n.a. & 1.56 $\pm$ 0.53*** \\
  & $\delta^{13}$C ($\permil$ V-PDB) & -29.84 $\pm$ 0.78 & -28.68 $\pm$ 0.51 & -29.46 $\pm$ 1.34 & -29.50 $\pm$ 1.01 & -31.94 $\pm$ 0.39 & -30.52 $\pm$ 0.35 & -31.22 $\pm$ n.a. & -31.66 $\pm$ 1.76 \\

@@

  \multicolumn{10}{l}{\textbf{Relative contents from mixing model (\%)}} \\
  & Carbohydrate$_{\text{NMR}}$ & 53.68 $\pm$ 7.07$^{\bullet}$ & 57.69 $\pm$ 2.31** & 42.67 $\pm$ 8.17$^{\bullet}$ & 44.60 $\pm$ 8.41** & 47.38 $\pm$ 4.89 & 63.13 $\pm$ 5.09 & 49.78 $\pm$ n.a. & 37.17 $\pm$ n.a. \\
  & Protein$_{\text{NMR}}$ & 10.76 $\pm$ 3.47** & 7.46 $\pm$ 1.36 & 19.81 $\pm$ 6.16** & 7.39 $\pm$ 2.22 & 7.48 $\pm$ 0.69 & 4.57 $\pm$ 0.35 & 9.36 $\pm$ n.a. & 7.39 $\pm$ n.a. \\
  & Lignin$_{\text{NMR}}$ & 21.05 $\pm$ 2.04 & 22.30 $\pm$ 0.98** & 25.10 $\pm$ 4.79 & 37.11 $\pm$ 4.57*** & 25.52 $\pm$ 2.46 & 20.54 $\pm$ 1.93 & 25.62 $\pm$ n.a. & 42.83 $\pm$ n.a. \\
  & Lipid$_{\text{NMR}}$ & 8.04 $\pm$ 1.28 & 5.93 $\pm$ 0.89 & 9.37 $\pm$ 1.85 & 3.42 $\pm$ 0.89 & 7.97 $\pm$ 0.92 & 4.63 $\pm$ 0.90 & 8.76 $\pm$ n.a. & 1.66 $\pm$ n.a. \\
  & Carbonyl$_{\text{NMR}}$ & 6.46 $\pm$ 2.99 & 6.62 $\pm$ 0.98** & 3.04 $\pm$ 2.41 & 7.49 $\pm$ 3.01** & 11.66 $\pm$ 2.18 & 7.13 $\pm$ 2.16 & 6.48 $\pm$ n.a. & 10.94 $\pm$ n.a. \\
  & Lignin:N & 132.24 $\pm$ 34.21 & 207.66 $\pm$ 36.13$^{\bullet}$ & 106.31 $\pm$ 31.66 & 351.58 $\pm$ 80.547$^{\bullet}$ & 275.88 $\pm$ 41.97 & 295.67 $\pm$ 14.96 & 236.74 $\pm$ n.a. & 428.00 $\pm$ n.a. \\

@@

  \multicolumn{10}{l}{\textbf{Lignin monomers}} \\
  & VSC (mg g$^{-1}$ C) & 58.89 $\pm$ 32.17 & 121.03 $\pm$ 47.54 & 51.96 $\pm$ 20.66 & 205.80 $\pm$ 81.72 & 39.37 $\pm$ 10.21 & 89.12 $\pm$ 40.70 & 40.41 $\pm$ n.a. & n.a. $\pm$ n.a. \\
  & S:V & 0.44 $\pm$ 0.10 & 0.66 $\pm$ 0.12** & 0.42 $\pm$ 0.19 & 1.56 $\pm$ 0.41** & 0.29 $\pm$ 0.06 & 0.77 $\pm$ 0.30 & 0.42 $\pm$ n.a. & n.a. $\pm$ n.a. \\
  & C:V & 0.19 $\pm$ 0.06 & 0.24 $\pm$ 0.03** & 0.24 $\pm$ 0.14 & 0.09 $\pm$ 0.01*** & 0.20 $\pm$ 0.14 & 0.18 $\pm$ 0.12 & 0.26 $\pm$ n.a. & n.a. $\pm$ n.a. \\
  & Ac:Al & 0.34 $\pm$ 0.08 & 0.32 $\pm$ 0.03* & 0.50 $\pm$ 0.23 & 0.55 $\pm$ 0.04* & 0.39 $\pm$ 0.27 & 0.22 $\pm$ 0.15 & 0.48 $\pm$ n.a. & n.a. $\pm$ n.a. \\
  & Ac:Al$_V$ & 0.11 $\pm$ 0.03 & 0.10 $\pm$ 0.02*** & 0.18 $\pm$ 0.04 & 0.87 $\pm$ 0.23*** & 0.11 $\pm$ 0.07 & 0.07 $\pm$ 0.03 & 0.15 $\pm$ n.a. & n.a. $\pm$ n.a. \\

@@

  \end{sidewaystable}

@@

The soil material clearly impacted the growth of the above- (plant) and belowground (root) biomass, with lower plant and root biomass and lower C and N contents in the subsoil compared to the topsoil (\ref{tab:M5-T1}). The soil material also influenced the elemental composition of belowground biomass, with broader C:N ratios in roots growing in the subsoil (\num{54.54 \pm 3.99}) compared to the topsoil (\num{34.43 \pm 6.69}; \(P < 0.001\)). The difference in root litter quality between the growing materials was further underlined by the higher lignin\textsubscript{NMR}:N ratio (\num{295.67 \pm 14.96} in subsoil compared to \num{207.66 \pm 36.13} in topsoil; \(P= 0.001\)). The decomposition of aboveground biomass was more progressed compared to belowground biomass, with approximately \SI{7}{\percent} of the biomass remaining after the detritusphere phase in both soil materials (compared to \(\sim\)\SI{22}{\percent} and \SI{13}{\percent} of roots remaining in topsoil and subsoil, respectively; \ref{tab:M5-T1}).

The molecular mixing model according to \citet{Nelson2005} showed how the composition of both above- and belowground biomass changed with decomposition, with the concentration of lignin\textsubscript{NMR} increasing from \SI{22.30 \pm 0.96}{\percent} to \SI{37.11 \pm 4.57}{\percent} in decomposing roots in the topsoil (\(P < 0.001\); \ref{tab:M5-T1}). This was accompanied by a slight decrease in carbohydrate\textsubscript{NMR} (from \SI{57.59 \pm 2.31}{\percent} to \SI{44.60 \pm 8.41}{\percent}; \(P= 0.008\)). These patterns differed from the aboveground biomass, where instead the relative protein\textsubscript{NMR} content almost doubled with decomposition (\(P=0.004\)), while the lignin content remained unaffected. The chemical composition of the decaying biomass in the subsoil could not be determined due to insufficient sample material remaining after decomposition.

The increased lignin\textsubscript{NMR} in decaying root biomass in the topsoil was corroborated by a slight, however not significant, increase in lignin monomers (from \SI{121.03 \pm 47.54}{mg\,VSC\,\gram^{-1}\,C} in living roots to \SI{205.80 \pm 81.72}{mg\,VSC\,\gram^{-1}\,C} in decaying roots; \(P=0.24\)). The progressed decomposition of the roots in the topsoil was reflected in an increased (Ac/Al)\textsubscript{V} ratio (from \num{0.10 \pm 0.02} to \num{0.87 \pm 0.23}; \citep{KoegelKnabner1988}) and overall, in changed lignin composition as indicated by the changing S:V, C:V and Ac:Al ratios. This was not reflected in the aboveground biomass (\ref{tab:M5-T1}).

@@

The changes in the oPOM<\SI{20}{\micro\metre} fractions were not observed in the subsoil. Instead, there was a slight (non-significant) increase in the C contribution of the fPOM fraction in the subsoil rhizosphere (from \SI{3.63 \pm 0.74}{\percent} to \SI{6.59 \pm 0.27}{\percent}; \(P=0.154\); \ref{fig:M5-F2}c). This led to increased C:N ratios of the fPOM fractions around living but also around decaying roots, from \num{24.37} to \num{26.88} in the rhizosphere (\(P=0.053\)) and from \num{22.91} to \num{25.36} in the detritusphere (\(P=0.086\)).

>>>>

    \vspace{0.5cm}
    \begin{justify}
    Kristina Witzgall\textsuperscript{1}, Franziska Steiner\textsuperscript{1}, Benjamin D. Hesse\textsuperscript{2,3}, Nicolás Riveras-Muñoz\textsuperscript{4}, Victoria Rodríguez\textsuperscript{5}, Pedro Teixeira\textsuperscript{1}, Li, M.\textsuperscript{6}, Rómulo Oses\textsuperscript{7}, Oscar Seguel\textsuperscript{8}, Steffen Seitz\textsuperscript{4}, Dirk Wagner\textsuperscript{5,6}, Thomas Scholten\textsuperscript{4}, Franz Buegger\textsuperscript{9} Gerrit Angst\textsuperscript{7,11,12}, and Carsten W. Mueller\textsuperscript{1,13,14}
    \end{justify}
    \vspace{0.2cm}
  \end{center}

@@

A critical aspect of the plant-associated contribution to these soil processes occurs at the root-soil interface, i.e., within the rhizosphere. As plants grow, their roots physically alter soil minerals, and root-associated microorganisms release weathering agents (e.g., organic acids; \citeauthor{Uroz2009}, \citeyear{Uroz2009}) that mobilize essential soil nutrients \citep{Gregory2022}. 
In addition, the release of rhizodeposits at the root-soil interface represents a major pathway for the build-up of rather persistent mineral-associated OM (MAOM) in soil \citep{Sokol2019, Villarino2021}. Moreover, the increased abundance of labile C in the rhizosphere stimulates the growth and activity of microorganisms. This in turn drives the formation of aggregates via microbially derived binding agents (i.e., extracellular polymeric substances; EPS) or via the enmeshment facilitated by fungal hyphae \citep{Chenu2011, Costa2018}. In addition, root exudates (e.g., root mucilage or polysaccharide exudates) can either adhere directly to mineral surfaces or function as glue-like binding agents, binding soil particles together, forming MAOM \citep{Roetzer2023} and driving soil aggregation \citep{Baumert2018}. As such, roots substantially shape the chemical, physical, and biological nature of the soil at the root-soil interface, and the rhizosphere is widely recognized as a dynamic system in both space and time \citep{Jones2009, Hinsinger2009, Kuzyakov2019}. However, relatively little is known about the temporal changes in processes occurring in the rhizosphere, as well as potential long-term effects of the living root system that may persist beyond the lifetime of the roots \citep{Oliver2021}. These effects, so-called 'root legacy effects', might have implications for soil physical, chemical, and biological parameters that can remain in the soil after plant death and thus shape the subsequent detritusphere \citep{Wurst2015}.

@@

Soil was sampled from five subplots on south-facing top slopes within the Santa Gracia Natural Reserve, located in the Coquimbo Region in Chile (\ang{29;44;26.3}S, \ang{71;09;01.2}W) with mean annual precipitation of \SI{89}{\milli\metre\,\year^{-1}} and mean annual air temperature of \SI{13.7}{\degreeCelsius} \citep{MinisterioObrasPublicas2017}. All five subplots were classified as Cambisol according to the WRB \citep{IUSS2015}. Vegetation was dominated by shrubs (\SIrange{30}{40}{\percent} coverage; including \textit{Proustia cuneifolia}, \textit{Balbisia peduncularis}, \textit{Cordia decandra} and \textit{Baccharis paniculatum}), classified as 'Interior Mediterranean desert scrub' \citep{Luebert2006}, and heavily disturbed by grazing animals \citep{Oeser2018}. The sampling site was selected as a suitable representation of an initial dryland soil with low vegetation cover, low organic C content, shallow and weakly structured A horizon \citep{Bernhard2018}. Topsoil and subsoil material was collected and sieved (\SI{< 2}{\milli\metre}) from the profile walls of A and B horizons (\SIrange{0}{3}{\centi\metre} and \SIrange{5}{30}{\centi\metre}, respectively), representing soils at different stages of development. This allowed us to directly compare topsoil and subsoil, but also to use the subsoil as a model system for initial soil formation via plant-soil interactions in a saprolitic material.

@@

The topsoil and subsoil material from the five subplots was homogenized and cleared of plant residues and seeds to prevent unwanted germination during incubation. The two soil materials (topsoil and subsoil) were weighed separately into pots (\SI{350}{\centi\metre\cubed}) aiming to resemble bulk density according to field conditions (\SIrange{1.6}{1.7}{\gram\per\centi\metre\cubed}; \citeauthor{Bernhard2018}, \citeyear{Bernhard2018}). Pre-germinated seeds (\textit{Helenium aromaticum} (Hook.) L. H. Bailey) were sown on half of the pots, while the other half was left free of plants and hereafter referred to as controls. After the initial incubation period of 70 days under greenhouse conditions ('rhizosphere phase'), half of the samples were harvested, while the remaining half were placed in complete darkness and incubated at room temperature for 100 days ('detritusphere phase'). The samples were left completely undisturbed during decomposition, meaning that above-ground biomass was left to decompose on the soil surface. During both incubation phases, the samples received \SI{6}{\milli\liter} of rainwater per day, without correcting for potential differences in soil moisture, e.g. due to differences in plant water use between treatments.

@@

We tested all parameters for homoscedasticity using Levene's test and the residuals of each model were tested for Normality using QQ-Plots and the Shapiro-Wilk test. In cases where the assumption of normality or homoscedasticity was not met, the data was appropriately transformed. To evaluate differences between the data presented in Table \ref{tab:M5-T1}, we conducted a one-way analysis of variance (ANOVA) with Tukey's honestly significant differences (HSD) as a post-hoc test. The remaining data was analyzed using mixed-effect models (package: \texttt{nlme}) using soil material (topsoil vs subsoil) and phase (rhizosphere phase vs detritusphere phase) as fixed factors and the ID as a random effect. All statistical testing was conducted in R (version 4.2.1; \citeauthor{RCoreTeam2008}, \citeyear{RCoreTeam2008}) using RStudio (Version 2022.07.2, R Core Team, Vienna, Austria; \citeauthor{RStudioTeam2015}, \citeyear{RStudioTeam2015}). Further, data shown in text and tables are given as the mean values and standard deviations.

\begin{table}[htbp]

@@

  \begin{tabular}{llcccc}

@@

  \textbf{Category} & \textbf{Property} & Rhizo Above & Rhizo Below & Detri Above & Detri Below \\

@@

  \multicolumn{6}{c}{\textbf{Topsoil}} \\

@@

  \multirow{8}{*}{Basic properties} & C (mg g$^{-1}$) & 4.06 $\pm$ 0.20 & 3.70 $\pm$ 0.08 & 3.49 $\pm$ 0.47 & 3.97 $\pm$ 0.74 \\
  & N (mg g$^{-1}$) & 0.17 $\pm$ 0.05** & 0.11 $\pm$ 0.02** & 0.25 $\pm$ 0.06** & 0.11 $\pm$ 0.02** \\
  & C:N & 25.61 $\pm$ 7.77** & 34.43 $\pm$ 3.43** & 14.67 $\pm$ 3.99*** & 37.26 $\pm$ 8.89 \\
  & Biomass (g) & 0.47 $\pm$ 0.17*** & 0.44 $\pm$ 0.13*** & 0.11 $\pm$ 0.07*** & 0.03 $\pm$ 0.02*** \\
  & Total biomass (g) & 0.90 $\pm$ 0.22*** & 0.13 $\pm$ 0.09*** & 0.54 $\pm$ 0.11*** & 0.05 $\pm$ 0.07*** \\
  & Root:shoot & 1.07 $\pm$ 0.48** & 0.32 $\pm$ 0.20** & 1.27 $\pm$ 0.37 & 0.24 $\pm$ n.a. \\
  & C (mg) & 1.91 $\pm$ 0.78*** & 1.62 $\pm$ 0.51*** & 0.38 $\pm$ 0.29*** & 0.12 $\pm$ 0.09*** \\
  & N (mg) & 0.07 $\pm$ 0.02*** & 0.05 $\pm$ 0.01*** & 0.03 $\pm$ 0.02*** & 0.00 $\pm$ 0.01*** \\
  & $\delta^{15}$N ($\permil$ air N$_2$) & 4.96 $\pm$ 0.78 & 4.14 $\pm$ 0.59$^{\bullet}$ & 4.29 $\pm$ 0.71 & 3.22 $\pm$ 0.81$^{\bullet}$ \\
  & $\delta^{13}$C ($\permil$ V-PDB) & -29.84 $\pm$ 0.78 & -28.68 $\pm$ 0.51 & -29.46 $\pm$ 1.34 & -29.50 $\pm$ 1.01 \\

@@

  \textbf{Relative contents} & Carbohydrate$_{\text{NMR}}$ & 53.68 $\pm$ 7.07$^{\bullet}$ & 57.69 $\pm$ 2.31** & 42.67 $\pm$ 8.17$^{\bullet}$ & 44.60 $\pm$ 8.41** \\
  & Protein$_{\text{NMR}}$ & 10.76 $\pm$ 3.47** & 7.46 $\pm$ 1.36 & 19.81 $\pm$ 6.16** & 7.39 $\pm$ 2.22 \\
  & Lignin$_{\text{NMR}}$ & 21.05 $\pm$ 2.04 & 22.30 $\pm$ 0.98** & 25.10 $\pm$ 4.79 & 37.11 $\pm$ 4.57*** \\
  & Lipid$_{\text{NMR}}$ & 8.04 $\pm$ 1.28 & 5.93 $\pm$ 0.89 & 9.37 $\pm$ 1.85 & 3.42 $\pm$ 0.89 \\
  & Carbonyl$_{\text{NMR}}$ & 6.46 $\pm$ 2.99 & 6.62 $\pm$ 0.98** & 3.04 $\pm$ 2.41 & 7.49 $\pm$ 3.01** \\
  & Lignin:N & 132.24 $\pm$ 34.21 & 207.66 $\pm$ 36.13$^{\bullet}$ & 106.31 $\pm$ 31.66 & 351.58 $\pm$ 80.547$^{\bullet}$ \\
  \textbf{Lignin monomers} & VSC (mg g$^{-1}$ C) & 58.89 $\pm$ 32.17 & 121.03 $\pm$ 47.54 & 51.96 $\pm$ 20.66 & 205.80 $\pm$ 81.72 \\
  & S:V & 0.44 $\pm$ 0.10 & 0.66 $\pm$ 0.12** & 0.42 $\pm$ 0.19 & 1.56 $\pm$ 0.41** \\
  & C:V & 0.19 $\pm$ 0.06 & 0.24 $\pm$ 0.03** & 0.24 $\pm$ 0.14 & 0.09 $\pm$ 0.01*** \\
  & Ac:Al & 0.34 $\pm$ 0.08 & 0.32 $\pm$ 0.03* & 0.50 $\pm$ 0.23 & 0.55 $\pm$ 0.04* \\
  & Ac:Al$_V$ & 0.11 $\pm$ 0.03 & 0.10 $\pm$ 0.02*** & 0.18 $\pm$ 0.04 & 0.87 $\pm$ 0.23*** \\
  \multicolumn{6}{c}{\textbf{Subsoil}} \\
  \multirow{8}{*}{Basic properties} & C (mg g$^{-1}$) & 3.16 $\pm$ 0.50 & 3.70 $\pm$ 0.09 & 3.23 $\pm$ n.a. & 3.36 $\pm$ 1.43 \\
  & N (mg g$^{-1}$) & 0.09 $\pm$ 0.01 & 0.07 $\pm$ 0.00 & 0.11 $\pm$ n.a. & 0.10 $\pm$ 0.04 \\
  & C:N & 33.58 $\pm$ 2.90 & 53.54 $\pm$ 5.27*** & 28.70 $\pm$ n.a. & 32.37 $\pm$ -- \\
  & Biomass (g) & 0.25 $\pm$ 0.09*** & 0.29 $\pm$ 0.07*** & 0.04 $\pm$ n.a. & 0.02 $\pm$ 0.00*** \\
  & C (mg) & 0.75 $\pm$ 0.21 & 1.07 $\pm$ 0.12*** & 0.36 $\pm$ n.a. & 0.06 $\pm$ 0.04*** \\
  & N (mg) & 0.02 $\pm$ 0.01 & 0.02 $\pm$ 0.00 & 0.01 $\pm$ n.a. & 0.00 $\pm$ 0.00 \\
  & $\delta^{15}$N ($\permil$ air N$_2$) & 4.66 $\pm$ 0.61 & 4.41 $\pm$ 0.45*** & 3.94 $\pm$ n.a. & 1.56 $\pm$ 0.53*** \\
  & $\delta^{13}$C ($\permil$ V-PDB) & -31.94 $\pm$ 0.39 & -30.52 $\pm$ 0.35 & -31.22 $\pm$ n.a. & -31.66 $\pm$ 1.76 \\
  \textbf{Relative contents} & Carbohydrate$_{\text{NMR}}$ & 47.38 $\pm$ 4.89 & 63.13 $\pm$ 5.09 & 49.78 $\pm$ n.a. & 37.17 $\pm$ n.a. \\
  & Protein$_{\text{NMR}}$ & 7.48 $\pm$ 0.69 & 4.57 $\pm$ 0.35 & 9.36 $\pm$ n.a. & 7.39 $\pm$ n.a. \\
  & Lignin$_{\text{NMR}}$ & 25.52 $\pm$ 2.46 & 20.54 $\pm$ 1.93 & 25.62 $\pm$ n.a. & 42.83 $\pm$ n.a. \\
  & Lipid$_{\text{NMR}}$ & 7.97 $\pm$ 0.92 & 4.63 $\pm$ 0.90 & 8.76 $\pm$ n.a. & 1.66 $\pm$ n.a. \\
  & Carbonyl$_{\text{NMR}}$ & 11.66 $\pm$ 2.18 & 7.13 $\pm$ 2.16 & 6.48 $\pm$ n.a. & 10.94 $\pm$ n.a. \\
  & Lignin:N & 275.88 $\pm$ 41.97 & 295.67 $\pm$ 14.96 & 236.74 $\pm$ n.a. & 428.00 $\pm$ n.a. \\
  \textbf{Lignin monomers} & VSC (mg g$^{-1}$ C) & 39.37 $\pm$ 10.21 & 89.12 $\pm$ 40.70 & 40.41 $\pm$ n.a. & n.a. $\pm$ n.a. \\
  & S:V & 0.29 $\pm$ 0.06 & 0.77 $\pm$ 0.30 & 0.42 $\pm$ n.a. & n.a. $\pm$ n.a. \\
  & C:V & 0.20 $\pm$ 0.14 & 0.18 $\pm$ 0.12 & 0.26 $\pm$ n.a. & n.a. $\pm$ n.a. \\
  & Ac:Al & 0.39 $\pm$ 0.27 & 0.22 $\pm$ 0.15 & 0.48 $\pm$ n.a. & n.a. $\pm$ n.a. \\
  & Ac:Al$_V$ & 0.11 $\pm$ 0.07 & 0.07 $\pm$ 0.03 & 0.15 $\pm$ n.a. & n.a. $\pm$ n.a. \\

@@

  \end{table}

@@

The soil material clearly impacted the growth of the above- (plant) and belowground (root) biomass, with lower plant and root biomass and lower C and N contents in the subsoil compared to the topsoil (Table \ref{tab:M5-T1}). The soil material also influenced the elemental composition of belowground biomass, with broader C:N ratios in roots growing in the subsoil (\num{54.54 \pm 3.99}) compared to the topsoil (\num{34.43 \pm 6.69}; \(P < 0.001\)). The difference in root litter quality between the growing materials was further underlined by the higher lignin\textsubscript{NMR}:N ratio (\num{295.67 \pm 14.96} in subsoil compared to \num{207.66 \pm 36.13} in topsoil; \(P= 0.001\)). The decomposition of aboveground biomass was more progressed compared to belowground biomass, with approximately \SI{7}{\percent} of the biomass remaining after the detritusphere phase in both soil materials (compared to \(\sim\)\SI{22}{\percent} and \SI{13}{\percent} of roots remaining in topsoil and subsoil, respectively; Table \ref{tab:M5-T1}).

The molecular mixing model according to \citet{Nelson2005} showed how the composition of both above- and belowground biomass changed with decomposition, with the concentration of lignin\textsubscript{NMR} increasing from \SI{22.30 \pm 0.96}{\percent} to \SI{37.11 \pm 4.57}{\percent} in decomposing roots in the topsoil (\(P < 0.001\); Table \ref{tab:M5-T1}). This was accompanied by a slight decrease in carbohydrate\textsubscript{NMR} (from \SI{57.59 \pm 2.31}{\percent} to \SI{44.60 \pm 8.41}{\percent}; \(P= 0.008\)). These patterns differed from the aboveground biomass, where instead the relative protein\textsubscript{NMR} content almost doubled with decomposition (\(P=0.004\)), while the lignin content remained unaffected. The chemical composition of the decaying biomass in the subsoil could not be determined due to insufficient sample material remaining after decomposition.

The increased lignin\textsubscript{NMR} in decaying root biomass in the topsoil was corroborated by a slight, however not significant, increase in lignin monomers (from \SI{121.03 \pm 47.54}{mg\,VSC\,\gram^{-1}\,C} in living roots to \SI{205.80 \pm 81.72}{mg\,VSC\,\gram^{-1}\,C} in decaying roots; \(P=0.24\)). The progressed decomposition of the roots in the topsoil was reflected in an increased (Ac/Al)\textsubscript{V} ratio (from \num{0.10 \pm 0.02} to \num{0.87 \pm 0.23}; \citeauthor{KoegelKnabner1988}, \citeyear{KoegelKnabner1988}) and overall, in changed lignin composition as indicated by the changing S:V, C:V and Ac:Al ratios. This was not reflected in the aboveground biomass (Table \ref{tab:M5-T1}).

@@

The changes in the oPOM<\SI{20}{\micro\metre} fractions were not observed in the subsoil. Instead, there was a slight (non-significant) increase in the C contribution of the fPOM fraction in the subsoil rhizosphere (from \SI{3.63 \pm 0.74}{\percent} to \SI{6.59 \pm 0.27}{\percent}; \(P=0.154\); Figure \ref{fig:M5-F2}c). This led to increased C:N ratios of the fPOM fractions around living but also around decaying roots, from \num{24.37} to \num{26.88} in the rhizosphere (\(P=0.053\)) and from \num{22.91} to \num{25.36} in the detritusphere (\(P=0.086\)).
